---
title: "Exploration"
output: 
  html_document: 
    fig_caption: yes
    toc: yes
    toc_float: true
---

<style type="text/css">

h1.title {
  font-size: 38px;
  color: #232D4B;
}
h1 { /* Header 1 */
  font-size: 28px;
  color: #E57200;
}
h2 { /* Header 2 */
    font-size: 22px;
  color: #666666;
}
h3 { /* Header 3 */
  font-size: 18px;
  font-family: "Times New Roman", Times, serif;
  color: #141E3C;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Some exploration for fun

With a focus on the legacy newspapers

## Volume and complexity of news

```{r vol, message = FALSE, cache = TRUE}
rm(list=ls())
library(tidyverse)
library(quanteda)
library(scales)
library(wordcloud)
library(RColorBrewer)

# Load the data and environment from readNews.R
setwd("~/Box Sync/mpc/dataForDemocracy/presidency_project/newspaper/")
load("workspaceR/newspaperExplore.RData")

date.vec <- seq(from=as.Date("2017-01-20"), to=as.Date("2018-03-02"), by="2 weeks") # update to Friday after last story
ggplot(qmeta2, aes(x=date)) + geom_point(stat="count") + 
  facet_wrap(~pub) +
  scale_x_date(labels = date_format("%m/%d"), breaks=date.vec) + 
  labs(title = "Volume of Trump Coverage", 
       subtitle = "New York Times, Washington Post, Wall Street Journal", 
       x="Date", y="Number of Stories") +
  theme(axis.text.x=element_text(angle=90, size=8), 
        plot.title = element_text(face="bold", size=18, hjust=0), 
        axis.title = element_text(face="bold", size=16),
        panel.grid.minor = element_blank(),
        legend.position="none")

qmeta2 %>% filter(readability<100) %>% 
  ggplot(aes(x = date, y = readability)) + 
  geom_jitter(aes(color=pub), alpha=0.05, width=0.25, height=0.0, size=2) +
  geom_smooth(aes(color=pub)) +
  scale_x_date(labels = date_format("%m/%d"), breaks=date.vec) + 
  ggtitle("'Readability' of Newspaper Coverage of Trump") +
  labs(y = "Readability (grade level)", x = "Date of Article") +
  scale_color_manual(values=c("blue3","turquoise","orange3"), name="Source") +
  theme(plot.title = element_text(face="bold", size=18, hjust=0),
        axis.title = element_text(face="bold", size=16),
        panel.grid.minor = element_blank(), legend.position = c(0.95,0.9),
        axis.text.x = element_text(angle=90),
        legend.text=element_text(size=10))

```

Via Flesch-Kincaid readability measure: $$readability = 0.39 \left(\frac{words}{sentences}\right) + 11.8\left(\frac{syllables}{words}\right) $$

<div style="margin-bottom:50px;">
</div>

## Sentiment/polarity of information and news

Via [sentimentr](https://cran.r-project.org/web/packages/sentimentr/README.html)

```{r tone, message = FALSE, cache = TRUE}
rm(list=ls())
library(tidyverse)
library(quanteda)
library(scales)

# Load the data and environment from readNews.R
setwd("~/Box Sync/mpc/dataForDemocracy/presidency_project/newspaper/")
load("workspaceR/newspaperSentiment.RData")

date.vec <- seq(from=as.Date("2017-01-20"), to=as.Date("2018-03-02"), by="2 weeks") # update to Friday after last story
# Plot!
ggplot(qmeta2, aes(x=date, y=avgSent)) +
  geom_jitter(aes(color=pub), width=0.2, height=0.0, size=2, alpha=.05) +
  geom_hline(yintercept=median(qmeta2$avgSent), color="gray50") +
  geom_smooth(aes(color=pub)) +
  scale_x_date(labels = date_format("%m/%d"), breaks=date.vec) +
  labs(title = "Sentiment of Trump Coverage",
       subtitle = "New York Times, Washington Post, Wall Street Journal", 
       y = "Average Sentiment across Sentences", x = "Date of Article") +
  scale_color_manual(values=c("blue3", "turquoise", "orange3"), name="Source") +
  theme(plot.title = element_text(face="bold", size=20, hjust=0),
        axis.title = element_text(face="bold", size=14),
        panel.grid.minor = element_blank(), legend.position = c(0.95,0.9),
        axis.text.x = element_text(angle=90),
        legend.text=element_text(size=12))

# Just opeds again
ggplot(filter(qmeta2, oped==1), aes(x=date, y=avgSent)) +
  geom_jitter(aes(color=pub), width=0.2, height=0.0, size=2, alpha=.15) +
  geom_hline(yintercept=median(qmeta2$avgSent), color="gray50") +
  geom_smooth(aes(color=pub)) +
  scale_x_date(labels = date_format("%m/%d"), breaks=date.vec) +
  labs(title = "Sentiment of Trump Op-Ed Coverage",
       subtitle = "New York Times, Washington Post, Wall Street Journal", 
       y = "Average Sentiment across Sentences", x = "Date of Article") +
  scale_color_manual(values=c("blue3", "turquoise", "orange3"), name="Source") +
  theme(plot.title = element_text(face="bold", size=20, hjust=0),
        axis.title = element_text(face="bold", size=14),
        panel.grid.minor = element_blank(), legend.position = c(0.95,0.9),
        axis.text.x = element_text(angle=90),
        legend.text=element_text(size=12))

```

<div style="margin-bottom:50px;">
</div>

## Moral foundations in news

More on [moral foundations](http://moralfoundations.org/)

### Care-Harm dimension

Within news articles

```{r harm, message = FALSE, cache = TRUE}
rm(list=ls())
library(tidyverse)
library(quanteda)
library(scales)
library(gridExtra)

# Load the data and environment from readNews.R
setwd("~/Box Sync/mpc/dataForDemocracy/presidency_project/newspaper/")
load("workspaceR/newspaperMoral.RData")

# Care-Harm
date.vec <- seq(from=as.Date("2017-01-20"), to=as.Date("2018-03-02"), by="2 weeks") # update to Friday after last story
# Plot
ph1 <- ggplot(qmeta2, aes(x=date, y=HarmVirtue)) +
  geom_jitter(aes(color=pub), alpha=0.15, width=0.2, height=0.0, size=2) +
  geom_hline(yintercept=mean(qmeta2$HarmVirtue), color="gray50") +
  geom_smooth(aes(color=pub)) +
  scale_x_date(labels = date_format("%m/%d"), breaks=date.vec) +
  ggtitle("Care MF: Trump Coverage") +
  labs(y = "Care Language", x = "Date") +
  scale_color_manual(values=c("blue3", "turquoise", "orange3"), name="Source") +
  theme(plot.title = element_text(face="bold", size=16, hjust=0),
        axis.title = element_text(face="bold", size=12),
        panel.grid.minor = element_blank(), legend.position = c(0.95,0.9),
        axis.text.x = element_text(angle=90),
        legend.text=element_text(size=10))

ph2 <- ggplot(filter(qmeta2, HarmVice<100), aes(x=date, y=HarmVice)) + 
  geom_jitter(aes(color=pub), alpha=0.15, width=0.2, height=0.0, size=2) +
  geom_hline(yintercept=mean(qmeta2$HarmVice), color="gray50") +
  geom_smooth(aes(color=pub)) +
  scale_x_date(labels = date_format("%m/%d"), breaks=date.vec) +
  ggtitle("Harm MF: Trump Coverage") +
  labs(y = "Harm Language", x = "Date") +
  scale_color_manual(values=c("blue3", "turquoise", "orange3"), name="Source") +
  theme(plot.title = element_text(face="bold", size=16, hjust=0),
        axis.title = element_text(face="bold", size=12),
        panel.grid.minor = element_blank(), legend.position = c(0.95,0.9),
        axis.text.x = element_text(angle=90),
        legend.text=element_text(size=10))

grid.arrange(ph1, ph2, nrow = 1)

```

Within op-eds

```{r harmoped, message = FALSE, cache = TRUE}
rm(list=ls())
library(tidyverse)
library(quanteda)
library(scales)
library(gridExtra)

# Load the data and environment from readNews.R
setwd("~/Box Sync/mpc/dataForDemocracy/presidency_project/newspaper/")
load("workspaceR/newspaperMoral.RData")

# Just opeds
# Plot
phoped1 <- ggplot(filter(qmeta2, oped==1), aes(x=date, y=HarmVirtue)) +
  geom_jitter(aes(color=pub), alpha=0.15, width=0.2, height=0.0, size=2) +
  geom_hline(yintercept=mean(qmeta2$HarmVirtue), color="gray50") +
  geom_smooth(aes(color=pub)) +
  scale_x_date(labels = date_format("%m/%d"), breaks=date.vec) +
  ggtitle("Care MF: Trump Op-Eds") +
  labs(y = "Care Language", x = "Date") +
  scale_color_manual(values=c("blue3", "turquoise", "orange3"), name="Source") +
  theme(plot.title = element_text(face="bold", size=16, hjust=0),
        axis.title = element_text(face="bold", size=12),
        panel.grid.minor = element_blank(), legend.position = c(0.95,0.9),
        axis.text.x = element_text(angle=90),
        legend.text=element_text(size=10))

phoped2 <- ggplot(filter(qmeta2, HarmVice<100 & oped==1), aes(x=date, y=HarmVice)) + 
  geom_jitter(aes(color=pub), alpha=0.15, width=0.2, height=0.0, size=2) +
  geom_hline(yintercept=mean(qmeta2$HarmVice), color="gray50") +
  geom_smooth(aes(color=pub)) +
  scale_x_date(labels = date_format("%m/%d"), breaks=date.vec) +
  ggtitle("Harm MF: Trump Op-Eds") +
  labs(y = "Harm Language", x = "Date") +
  scale_color_manual(values=c("blue3", "turquoise", "orange3"), name="Source") +
  theme(plot.title = element_text(face="bold", size=16, hjust=0),
        axis.title = element_text(face="bold", size=12),
        panel.grid.minor = element_blank(), legend.position = c(0.95,0.9),
        axis.text.x = element_text(angle=90),
        legend.text=element_text(size=10))

grid.arrange(phoped1, phoped2, nrow = 1)
```

<div style="margin-bottom:50px;">
</div>

### Ingroup/Loyalty-Other/Betrayal dimension

Within news articles

```{r ingroup, message = FALSE, cache = TRUE}
rm(list=ls())
library(tidyverse)
library(quanteda)
library(scales)
library(gridExtra)

# Load the data and environment from readNews.R
setwd("~/Box Sync/mpc/dataForDemocracy/presidency_project/newspaper/")
load("workspaceR/newspaperMoral.RData")

# Ingroup/loyalty-Others/betrayal
pi1 <- ggplot(filter(qmeta2, IngroupVirtue<100), aes(x=date, y=IngroupVirtue)) + 
  geom_jitter(aes(color=pub), alpha=0.15, width=0.2, height=0.0, size=2) +
  geom_hline(yintercept=mean(qmeta2$FairnessVirtue), color="gray50") +
  geom_smooth(aes(color=pub)) +
  scale_x_date(labels = date_format("%m/%d"), breaks=date.vec) +
  ggtitle("Ingroup MF: Trump Coverage") +
  labs(y = "Ingroup Language", x = "Date") +
  scale_color_manual(values=c("blue3", "turquoise", "orange3"), name="Source") +
  theme(plot.title = element_text(face="bold", size=16, hjust=0),
        axis.title = element_text(face="bold", size=12),
        panel.grid.minor = element_blank(), legend.position = c(0.95,0.9),
        axis.text.x = element_text(angle=90),
        legend.text=element_text(size=10))

pi2 <- ggplot(filter(qmeta2, IngroupVice<100), aes(x=date, y=IngroupVice)) + 
  geom_jitter(aes(color=pub), alpha=0.15, width=0.2, height=0.0, size=2) +
  geom_hline(yintercept=mean(qmeta2$FairnessVice), color="gray50") +
  geom_smooth(aes(color=pub)) +
  scale_x_date(labels = date_format("%m/%d"), breaks=date.vec) +
  ggtitle("Outgroup MF: Trump Coverage") +
  labs(y = "Outgroup Language", x = "Date") +
  scale_color_manual(values=c("blue3", "turquoise", "orange3"), name="Source") +
  theme(plot.title = element_text(face="bold", size=16, hjust=0),
        axis.title = element_text(face="bold", size=12),
        panel.grid.minor = element_blank(), legend.position = c(0.95,0.9),
        axis.text.x = element_text(angle=90),
        legend.text=element_text(size=10))

grid.arrange(pi1, pi2, nrow = 1)
```

Within op-eds

```{r ingroupoped, message = FALSE, cache = TRUE}
rm(list=ls())
library(tidyverse)
library(quanteda)
library(scales)
library(gridExtra)

# Load the data and environment from readNews.R
setwd("~/Box Sync/mpc/dataForDemocracy/presidency_project/newspaper/")
load("workspaceR/newspaperMoral.RData")

# OPEDS: Ingroup/loyalty-Others/betrayal
pioped1 <- ggplot(filter(qmeta2, IngroupVirtue<100 & oped==1), aes(x=date, y=IngroupVirtue)) + 
  geom_jitter(aes(color=pub), alpha=0.15, width=0.2, height=0.0, size=2) +
  geom_hline(yintercept=mean(qmeta2$FairnessVirtue), color="gray50") +
  geom_smooth(aes(color=pub)) +
  scale_x_date(labels = date_format("%m/%d"), breaks=date.vec) +
  ggtitle("Ingroup MF: Trump Op-Eds") +
  labs(y = "Ingroup Language", x = "Date") +
  scale_color_manual(values=c("blue3", "turquoise", "orange3"), name="Source") +
  theme(plot.title = element_text(face="bold", size=16, hjust=0),
        axis.title = element_text(face="bold", size=12),
        panel.grid.minor = element_blank(), legend.position = c(0.95,0.9),
        axis.text.x = element_text(angle=90),
        legend.text=element_text(size=10))

pioped2 <- ggplot(filter(qmeta2, IngroupVice<100 & oped==1), aes(x=date, y=IngroupVice)) + 
  geom_jitter(aes(color=pub), alpha=0.15, width=0.2, height=0.0, size=2) +
  geom_hline(yintercept=mean(qmeta2$FairnessVice), color="gray50") +
  geom_smooth(aes(color=pub)) +
  scale_x_date(labels = date_format("%m/%d"), breaks=date.vec) +
  ggtitle("Outgroup MF: Trump Op-Eds") +
  labs(y = "Outgroup Language", x = "Date") +
  scale_color_manual(values=c("blue3", "turquoise", "orange3"), name="Source") +
  theme(plot.title = element_text(face="bold", size=16, hjust=0),
        axis.title = element_text(face="bold", size=12),
        panel.grid.minor = element_blank(), legend.position = c(0.95,0.9),
        axis.text.x = element_text(angle=90),
        legend.text=element_text(size=10))

grid.arrange(pioped1, pioped2, nrow = 1)
```

<div style="margin-bottom:50px;">
</div>

## Topics/themes in news

Via LDA topic model

```{r topic, message = FALSE, cache = TRUE}
rm(list=ls())
library(tidyverse)
library(quanteda)
library(topicmodels)
library(ggridges)
library(lubridate)

# Load the data and environment from readNews.R
setwd("~/Box Sync/mpc/dataForDemocracy/presidency_project/newspaper/")
load("workspaceR/newspaperTopicModel.RData")

ggplot(topicweeklong2, aes(x=week, y=terms2, height=prevalence, group=terms2)) + 
  ggtitle("Topical Prevalence by Week") +
  labs(y = "", x = "Weeks since Inauguration") +
  scale_x_continuous(breaks=seq(5,50,5)) +
  geom_density_ridges(stat="identity", scale = 2, fill="lightblue", color="lightblue") + 
  theme_ridges(font_size=6)

ggplot(topicpublong, aes(x=reorder(terms, prev), y=prev, fill=pub)) +
  geom_bar(stat="identity", position="dodge", width=0.5) +
  scale_fill_manual(values=c("blue3", "turquoise", "orange3")) + 
  labs(x="Topic", y="Prevalence Percent") +
  coord_flip() + ggtitle("Topic Prevalence by Publication") +
  theme(legend.text=element_text(size=6))


```

See more on topics [here](http://people.virginia.edu/~mpc8t/datafordemocracy/papertopics100_2018_03/).

# So much to do

## Analyze

 - issue emphases, ideological positioning, named entities, events, ...
 - combinations of features (polarity and issues, morality and topics)
 - word embeddings for "partisan bias" in news sources?
 
## Build

 - dynamic visualizations
 - digital, searchable resource
